---
title: "CSS & JavaScript Dependencies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CSS & JavaScript Dependencies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Overview

Often times a D3 visualization will require additional components such as CSS styles,  JavaScript helper functions, or even entire JavaScript libraries. This article describes how to include these dependencies along with your visualization.

## CSS Styles

It's common to require a few CSS styles when creating a visualization. By convention, if you include a CSS file with the same base filename as your D3 script file it will automatically be included. For example, consider a D3 script "barchart.js" within a sub-directory:

```bash
barchart.js
barchart.css
```

If you render the "barchart.js" script like this:

```{r}
r2d3(data=c(0.3, 0.6, 0.8, 0.95, 0.40, 0.20), script = "barchart.js")
```

Then the "barchart.css" file will be automatically included.

You can also name a CSS file "styles.css" (useful if you want multiple visualizations to share styles) and it will also be automatically included:

```bash
barchart.js
styles.css
```

Finally, you can explicitly include a CSS file of any name via the `css` parameter of `r2d3()`:

```{r}
r2d3(data=c(0.3, 0.6, 0.8, 0.95, 0.40, 0.20), script = "barchart.js", css = "theme.css")
```

## JavaScript

If your visualization depends on additional JavaScript files (e.g. some common code that you want to re-use) you can specify this using the `dependencies` parameter of `r2d3()`. For example:

```{r}
r2d3(
  data=c(0.3, 0.6, 0.8, 0.95, 0.40, 0.20), 
  script = "barchart.js", dependencies = "utils.js"
)
```

## External Libraries

You can include entire external libraries as dependencies by passing an "html_dependency" object created using the `htmltools::htmlDependency()` function. For example, here's how you would include the version of jQuery bundled with the **rmarkdown** package:

```{r}
r2d3(
  data=c(0.3, 0.6, 0.8, 0.95, 0.40, 0.20), 
  script = "barchart.js", dependencies = rmarkdown::html_dependency_jquery()
)
```

Note that you can pass a list of multiple dependencies (either paths to JavaScript files or "html_dependencies" objects) via the `dependencies` argument.





